### 2020-09-04 presentation/slides.Rmd
desactivate renv

---
title: "Urban and socio-economic correlates of property prices in Dublin's area"
subtitle: |
"The 7th IEEE International Conference on Data Science and Advanced Analytics"
"Special Session: Environmental and Geo-spatial Data Analytics (EnGeoData)"
author: "Damien Dupré"
date: "Dublin City University - October 6th, 2020"
output:
  xaringan::moon_reader:
    css: ["default", "metropolis", "metropolis-fonts", "css/custom_design.css"]
    lib_dir: libs
    nature:
      beforeInit: "libs/cols_macro.js"
      highlightStyle: zenburn
      highlightLines: true
      countIncrementalSlides: false
params:
  eval_rights: FALSE
---

```{r setup, include = FALSE}
# general options --------------------------------------------------------------
options(scipen = 999)
set.seed(123)
# chunk options ----------------------------------------------------------------
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE,
  cache = FALSE,
  comment = "", 
  fig.align = "center", 
  fig.retina = 3
  )
# libraries --------------------------------------------------------------------
library(tidyverse)
library(fontawesome)
```

layout: true

<div class="custom-footer"><span>DSAA EnGeoData - Dupré (2020) </span></div>

---

# My Journey into Data Science

#### Developement of the DynEmo Facial Expression Database (Master)

* Dynamic and spontaneous emotions
* Assessed with self-reports and by observers

#### Analysis of Emotional User Experience of Innovative Tech. (Industrial PhD)

* Understand users' acceptance of technologies from their emotional response
* Based on multivariate self-reports

#### Evaluation of Emotions from Facial and Physiological Measures (Industrial PostDoc)

* Applications to marketing, sports and automotive industries
* Dynamic changes with trend extraction techniques (2 patents)

#### Performance Prediction using Machine Learning (Academic PostDoc)

* Application to sport analytics
* Big Data treatment (> 1 million users with activities recorded in the past 5 years)

---
class: inverse, mline, center, middle

# 1. A Closer Look at Property Prices

---

# Market Characteristics

An housing market is endemic to a country in a specific context.

Countries are very different in term of:
- Styles
- Materials
- Legislations
- Urban features
- Population characteristics
#https://www.visualcapitalist.com/mapped-the-countries-with-the-highest-housing-bubble-risks/

![Plot title. ](slides_insertimage_1.png)

---
class: inverse, mline, left, middle

<img class="circle" src="https://github.com/damien-dupre.png" width="250px"/>

# Thanks for your attention, find me at...

[`r fa(name = "twitter")` @damien_dupre](http://twitter.com/damien_dupre)  
[`r fa(name = "github")` @damien-dupre](http://github.com/damien-dupre)  
[`r fa(name = "link")` damien-datasci-blog.netlify.app](https://damien-datasci-blog.netlify.app)  
[`r fa(name = "paper-plane")` damien.dupre@dcu.ie](mailto:damien.dupre@dcu.ie)